An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-09-30 nips_2016.tex
More touch-ups to Section 6.2.

In practice, it is difficult to compute multinomial probabilities for large $n$,
in that case the per-bin distributions are well-approximated by Poisson
distributions with parameter $\lambda_i=P_i n$. Given a set of samples $\Theta =
\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
$\theta_i$ that fall in bin $j$. Then the log probability for the set is
approximated as
\begin{equation}\label{eq:log_prob_poissons}
    \log p(\Theta) \approx \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
\end{equation}

The results for each model are shown in Table~\ref{tab:poissons} ...
{\color{blue} Daniel: IN PROGRESS}

@@

Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used
by the three methods in each iteration. Our method consumes significantly less
data; most minibatch sizes are smaller than 1000, and the average size is 210.
The other methods occasionally need to consume a large proportion of the
entire data set of one million elements; the average minibatch sizes are 14862
and 18098 for Korattikara 2014 and Bardenet 2014, respectively. The average
minibatch sizes accurately predict the running times of these methods since
all methods have a running time proportional to the total data consumed, with
one caveat: the method of~\cite{icml2014c1_bardenet14} requires a global bound
on the term $C_{\theta,\theta'}$ over the entire the data set at every iteration
in order to calculate the error bound (in Equation~\ref{eq:bad_bound} of this
paper, i.e., Equation 9 of~\cite{icml2014c1_bardenet14}).  This bound must be
either computed analytically using properties of the specific dataset, or it
must be checked at each sample if a black-box test is required.  We found that
the sample code provided by the authors of~\cite{icml2014c1_bardenet14} in fact
computed this bound explicitly for each sample generated, i.e. the
implementation traversed the entire dataset for each sample, providing no
performance advantage over the complete test. For this reason, we omitted the
time to compute $C_{\theta,\theta'}$ from our measurements of the running time
of the sample implementation of \cite{icml2014c1_bardenet14}, and consider only
the time to process the minibatches.

@@

\begin{table}[t]
    \caption{{\color{blue} Daniel: TODO Poisson Probabilities, also we want
    ground truth right?}}
    \label{tab:poissons}
    \centering
    \begin{tabular}{l l l l l}
    \toprule
    Model & Ground Truth & Ours & Korattikara 2014 & Bardenet 2014 \\
    \midrule
    Poission Value & TODO & TODO & TODO & TODO \\
    \bottomrule
    \end{tabular}
\end{table}

@@

Figure~\ref{fig:logistic_performance} shows the prediction accuracy and log
likelihood on the test set as a function of the cumulative training data points
processed.  We see that our minibatch MH test is more efficient; it has similar
or better performance while consuming fewer data points.

@@

Figure~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes for
all three methods. With an initial minibatch size of 163, Korattikara 2014 and
Bardenet 2014 achieve average minibatch size of 2025 and 7612 respectively over
the MNIST classification task, while our method achieves that of 100, showing
significant better performance than the benchmark test methods.

>>>>

In practice, it is difficult to compute multinomial probabilities for large $n$,
in that case the per-bin distributions are well-approximated by Poisson
distributions with parameter $\lambda_i=P_i n$. Given a set of samples $\Theta =
\{\theta_1,\ldots,\theta_T\}$, let $c_j$ denote the number of individual samples
$\theta_i$ that fall in bin $j$. Then the log probability for the set is
approximated as
\begin{equation}\label{eq:log_prob_poissons}
    \log p(\Theta) \approx \frac{1}{N_{\rm bins}} \sum_{j=1}^{N_{\rm bins}} c_j \log (n P_j) - n P_j - \log(\Gamma(c_j+1)).
\end{equation}
The results for each model are shown in Table~\ref{tab:poissons} ...
{\color{blue} Daniel: IN PROGRESS}

@@

Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used
by the three methods in each iteration. Our method consumes significantly less
data; most minibatch sizes are smaller than 1000, and the average size is 210.
The other methods occasionally need to consume a large proportion of the
entire data set of one million elements; the average minibatch sizes are 14862
and 18098 for Korattikara 2014 and Bardenet 2014, respectively. The average
minibatch sizes accurately predict the running times of these methods since
all methods have a running time proportional to the total data consumed, with
one caveat: the method of~\cite{icml2014c1_bardenet14} requires a global bound
on the term $C_{\theta,\theta'}$ (see Equation~\ref{eq:bad_bound}) over the
entire the data set at every iteration in order to calculate the error bound.
This bound must be either computed analytically using properties of the specific
dataset, or it must be checked at each sample if a black-box test is required.
We found that the sample code provided by the authors
of~\cite{icml2014c1_bardenet14} in fact computed this bound explicitly for each
sample generated, i.e. the implementation traversed the entire dataset for each
sample, providing no performance advantage over the complete test. For this
reason, we omitted the time to compute $C_{\theta,\theta'}$ from our
measurements of the running time of the sample implementation
of~\cite{icml2014c1_bardenet14}, and consider only the time to process the
minibatches.

@@

\begin{table}[t]
    \caption{\color{blue} Daniel: TODO Poisson Probabilities}
    \label{tab:poissons}
    \centering
    \begin{tabular}{l l l l}
    \toprule
    Model & Ours & Korattikara 2014 & Bardenet 2014 \\
    \midrule
    Poisson Value (Eq.~\ref{eq:log_prob_poissons}) & TODO & TODO & TODO \\
    \bottomrule
    \end{tabular}
\end{table}

@@

Figure~\ref{fig:logistic_performance} shows the prediction accuracy and log
likelihood on the test set as a function of the cumulative training data points
processed.  We see that our minibatch MH test is more efficient compared to the
other two methods, because it has similar or better performance while consuming
fewer data points.

@@

Figure~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes for
all three methods. With an initial minibatch size of 100, Korattikara 2014 and
Bardenet 2014 achieve average minibatch size of 2026 and 7612 respectively over
the MNIST classification task, while our method achieves that of 163, showing
significantly better performance than the benchmark test methods.
